---
title: "ricmd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ricmd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
```

# Introduction

The integrated Rule Oriented Data System (iRODS) is an open source
data management software. It is used by research organisations and
governmental institutes and has an important role in their ever
growing data services.

This package provides an R user interface with iRODS and implements
the basic functionality of working with iRODS collections and data
objects from R. Whith this package it is possible to create, modify,
and remove collections and data objects from an iRODS Zone. For data objects it
is also possible to add, remove and modify meta data in the so called
attribute-value-unit (AVU) tripples. For more information see the [iRODS
documentation](https://irods.org/documentation/)


This package is based on the [python
irodsclient](https://github.com/irods/python-irodsclient) and it calls
[reticulate](https://cran.r-project.org/package=reticulate) to interact
with this library.

All functions in this package start with the ri\_ prefix, this
prevents hiding functions from other packages and makes your code more
readable.


# prerequisites

This package requires that the python-irodsclient library is
installed. It is developped and tested using python version >3. Use
pip to install the library:

```{}
pip3 install python-irodsclient
```

This package uses the iRODS environment files in `~/.irods/`, these
files should contain the user credentials needed to connect to iRODS.
So it is up to the user to set up their iRODS environment before he or
she can use this package. If the user can use the so called 'i
commands', this package should also function properly.

# Zones, collections, data objects and meta data

To understand how to use this package a basic understanding of iRODS
is needed. The short version to remember is: each iRODS system is a
'zone', in these zones data collections exists and each data collection
contains many data objects. Each data object contains meta data in so
called Attribute-Value-Unit (AVU) tripples. There is a longer version
but that's beyond the scope of this vignette, please see the [iRODS
documentation](https://irods.org/documentation/).

## Zones

A Zone is an independent iRODS system which contains the data catalog
and one or more iRODS servers. While an iRODS implementation can
contain multiple zones, in this vignette we assume a single zone
implementation.

## Collections

A Zone can contain on or more collections and each collection can
contain one more sub collections. Collections are a way to store data
in a hierarchical structure like directories or folders.

This package uses a default collection, which you must set manually.
If you manipulate (store/retrieve etc) your data objects, it is assumed
that they exists in your default collection, unless you provide a specific
collection in the function arguments.


## Data objects

A Collection can contain multiple data objects, like files in a folder or
directory. Although somewhat different, a data object can be
considered as a normal file. 

## Meta data
 
Meta data is stored in AVU tripples. Many objects in iRODS, like
users, collections, or resources can contain meta data but this
package only considers meta data associated with data objects.

AVU triples must contain an attribute and value, unit is optional.
Within iRODS it is possible to search and query data objects based on
their values in the AVU triples. These search options are not yet
implemented in this package.


# Initialise an iRODS session


So, now that we know the basics let's start with an iRODS session. We
assume that your iRODS environment is set up properly and that you
have the python iRODS client in place.

To start an iRODS session from within R you exexcute the following
commands:

```{r}
library(ricmd)
ri_session()
```

Executing this command creates a hidden environment wich contains the
iRODS configuration and the session object. This session object refers
to the session class from the pyton library and contains all the methods
for working with iRODS.

After initiating the iRODS session one must set the default
collection:

```{r}
ri_setCollection("/tempZone/home/devel")
```

The name of the collection always start with the name of the zone, in
our case '/tempZone' (mind the '/' at the beginning). We use the
standard collection from a default iRODS installation with a username
'devel', hence 'home/devel'.

This default collection can be changed anytime in your script. If you
want check which collection is the default collection, then use:

```{r}
ri_getCollection()
```

# Creating a new collection

A new data collection can be created with

```{r}
ri_createCollection("/tmpZone/home/devel/tests")
```























 * Session
 * Collection
 * Data objects  
 * Files and data path  
 * Initialise  

# Create, remove and select collections

# Store, retrieve and remove data objects

# Manage meta data
